---
title: "Instacart Exploratory Analysis"
output:
  html_document: default
  html_notebook: default
---


```{r}
source("readDataToMemory.R")
readInstacart()

library(DBI)
library(ggplot2)
library(ggthemes)

src_tbls(sc)
```

### Order products

`order_products__*.csv`

These files specify which products were purchased in each order.

```{r}
dbGetQuery(sc, "SELECT * FROM order_products__prior_tbl LIMIT 10")
```
 
## orders

`orders.csv`

This file tells to which set (prior, train, test) an order belongs. 


```{r}
orders_prev <- dbGetQuery(sc, "SELECT * FROM orders_tbl LIMIT 100")
orders_prev
```
 
```{r}
dbGetQuery(
  sc,
  "
SELECT COUNT(DISTINCT product_id) AS products
FROM order_products__prior_tbl
  "
  )
```

```{r}
products_popularity_txt <- "
SELECT product_id
,   COUNT(1) AS n_orders
FROM order_products__prior_tbl
GROUP BY product_id
ORDER BY n_orders DESC
LIMIT 50
"

products_popularity <-
  dbGetQuery(sc, products_popularity_txt)

products_popularity
```

## Products

`products.csv`

```{r}
dbGetQuery(sc, "SELECT * FROM products_tbl LIMIT 20")
```

## Joins

```{r}
products_popularity_txt <- "
SELECT op.product_id
,   n_orders
,   product_name
FROM (
    SELECT product_id
    ,   COUNT(1) AS n_orders
    FROM order_products__prior_tbl
    GROUP BY product_id
    ORDER BY n_orders DESC
    LIMIT 30) op
LEFT JOIN (
    SELECT product_id
    ,   product_name
    FROM products_tbl) p
ON op.product_id = p.product_id
"

products_popularity <-
  dbGetQuery(sc, products_popularity_txt)

products_popularity
```

```{r}

products_popularity %>%
  ggplot(
    aes(reorder(product_name, n_orders, function(x) x), 
        n_orders)) +
  geom_bar(stat="identity", fill='darkorange') +
  coord_flip() + 
  scale_y_continuous(label=scales::comma) +
  xlab("product") +
  ylab("Number of orders") +
  theme_minimal()
```

```{r}
products_popularity_txt <- "
SELECT op.product_id
,   n_orders
,   product_name
FROM (
    SELECT product_id
    ,   COUNT(1) AS n_orders
    FROM order_products__prior_tbl
    GROUP BY product_id
    ORDER BY n_orders DESC
    LIMIT 30) op
LEFT JOIN (
    SELECT product_id
    ,   product_name
    FROM products_tbl) p
ON op.product_id = p.product_id
"

products_popularity <-
  dbGetQuery(sc, products_popularity_txt)

products_popularity
```

## Questions

* Which day of week people order most fequently?

```{r}
dow_freq_txt <- "
SELECT order_dow
,   COUNT(1) AS dow_f
FROM orders_tbl
GROUP BY order_dow
ORDER BY dow_f DESC
"

dow_frequency <-
  dbGetQuery(sc, dow_freq_txt)

dow_frequency
```

* The most popular hour of doing order? 

```{r}
hod_freq_txt <- "
SELECT order_hour_of_day
,   COUNT(1) AS hod_f
FROM orders_tbl
GROUP BY order_hour_of_day
ORDER BY hod_f DESC
"

hod_frequency <-
  dbGetQuery(sc, hod_freq_txt)

hod_frequency
```

* The most reordered products?

```{r}
most_reord_txt <- "
SELECT product_id
,   COUNT(reordered) AS reord
FROM order_products__prior_tbl
GROUP BY product_id
ORDER BY reord DESC
"

most_reord <-
  dbGetQuery(sc, most_reord_txt)

most_reord
```

* The less reordered product?

```{r}
least_reord_txt <- "
SELECT product_id
,   COUNT(reordered) AS reord
FROM order_products__prior_tbl
GROUP BY product_id
ORDER BY reord
"

least_reord <-
  dbGetQuery(sc, least_reord_txt)

least_reord
```

* Plot histogram of number of products in an order.

```{r}
n_txt <- "
SELECT order_id,
COUNT(1) AS n_ord
FROM order_products__prior_tbl
GROUP BY order_id
ORDER BY n_ord DESC
LIMIT 30
"

n <-
  dbGetQuery(sc, n_txt)

n
```

```{r}

n %>%
  ggplot(
    aes(reorder(order_id, n_ord, function(x) x), 
        n_ord)) +
  geom_bar(stat="identity", fill='steelblue') +
  coord_flip() + 
  scale_y_continuous(label=scales::comma) +
  xlab("order") +
  ylab("Number of products") +
  theme_minimal()
```


```{r}
dbGetQuery(sc,"
SELECT n_products,
COUNT(order_id) AS n_orders
FROM (
  SELECT order_id, COUNT(1) AS n_products
  FROM order_products__prior_tbl
  GROUP BY order_id) op
WHERE n_products <= 50
GROUP BY n_products
") %>% 
  ggplot(aes(n_products,n_orders))+
  geom_col(fill="steelblue")+
  theme_minimal()
```

```{r}
order_products__prior %>%
  group_by(order_id) %>%
  summarise(n_products=n())%>%
  group_by(n_products)%>%
  summarise(n_orders=n())%>%
  filter(n_products <= 50) %>%
  collect%>%
  ggplot(aes(n_products,n_orders))+
  geom_col(fill="steelblue")+
  theme_minimal()
```


* What is the most popular department? 

```{r}
dep_popularity_txt <- "
SELECT department_id,
COUNT(1) AS n_ord
FROM (
    SELECT product_id
    ,   COUNT(1) AS n_orders
    FROM order_products__prior_tbl
    GROUP BY product_id
    ORDER BY n_orders DESC) op
LEFT JOIN (
    SELECT product_id
    ,   department_id
    FROM products_tbl) p
ON op.product_id = p.product_id
GROUP BY department_id
ORDER BY n_ord DESC
"

dep_popularity <-
  dbGetQuery(sc, dep_popularity_txt)

dep_popularity
```

## Whisker, magrittr

```{r}
library(whisker)
library(magrittr)

product_times_ordered_txt <- "
SELECT n_orders
,   COUNT(DISTINCT user_id) AS n_users
FROM (
    SELECT o.user_id
    ,   COUNT(DISTINCT op.order_id) AS n_orders
    FROM (
        SELECT order_id
        FROM order_products__prior_tbl
        WHERE product_id = {{productId}}
    ) op
    LEFT JOIN (
        SELECT  user_id
        ,   order_id
        FROM orders_tbl) o
    ON op.order_id = o.order_id
    GROUP BY o.user_id) opo
GROUP BY n_orders
ORDER BY n_orders
"

data <- list(productId = 24852)

product_times_ordered_txt %>%
  whisker.render(data) %>%
  cat
```
```{r}
product_times_ordered_txt %>%
  whisker.render(data) %>%
  dbGetQuery(sc, .) %>%
  ggplot(aes(n_orders, n_users)) + 
  geom_line() +
  theme_minimal()
```

## Questions

* For given department, plot the most popular product (with templating department).

```{r}
products_popularity_txt <- "
SELECT op.product_id
,   n_orders
,   product_name
FROM (
    SELECT product_id
    ,   COUNT(1) AS n_orders
    FROM order_products__prior_tbl
    GROUP BY product_id
    ORDER BY n_orders DESC) op
RIGHT JOIN (
    SELECT product_id
    ,   product_name
    FROM products_tbl
    WHERE department_id = 19) p
ON op.product_id = p.product_id
LIMIT 30
"

products_popularity <-
  dbGetQuery(sc, products_popularity_txt)

products_popularity
```

```{r}

products_popularity %>%
  ggplot(
    aes(reorder(product_name, n_orders, function(x) x), 
        n_orders)) +
  geom_bar(stat="identity", fill='darkorange') +
  coord_flip() + 
  scale_y_continuous(label=scales::comma) +
  xlab("product") +
  ylab("Number of orders") +
  theme_minimal()
```

* For given day, rank department popularity.

```{r}
dep_popularity_txt <- "
SELECT department_id, COUNT(department_id) AS popularity
FROM (
  SELECT product_id
  FROM order_products__train_tbl AS a INNER JOIN
      orders_tbl AS b ON a.order_id=b.order_id
  WHERE order_dow=0) AS ab INNER JOIN
  products_tbl AS c
  ON ab.product_id = c.product_id
  GROUP BY department_id
  ORDER BY popularity DESC
"


dep_popularity <-
  dbGetQuery(sc, dep_popularity_txt)
```


```{r}

dep_popularity %>%
  ggplot(
    aes(reorder(department_id, popularity, function(x) x), 
        popularity)) +
  geom_bar(stat="identity", fill='darkorange') +
  coord_flip() + 
  scale_y_continuous(label=scales::comma) +
  xlab("department") +
  ylab("Popularity") +
  theme_minimal()
```

## More questions

* For a given department, give a list of the 10 most popular product (with templating department).



* For a given day of the week, rank department popularity.

```{r}
library(whisker)
library(magrittr)

dep_popularity_txt <- "
SELECT department_id,
COUNT(1) AS n_ord
FROM (
    SELECT product_id
    ,   COUNT(1) AS n_orders
    FROM order_products__prior_tbl AS opp
    INNER JOIN
      orders_tbl AS b ON opp.order_id=b.order_id
    WHERE order_dow={{day}}
    GROUP BY product_id
    ORDER BY n_orders DESC) AS op
LEFT JOIN (
    SELECT product_id
    ,   department_id
    FROM products_tbl) p
ON op.product_id = p.product_id
GROUP BY department_id
ORDER BY n_ord DESC
"

data <- list(day = 2)

dep_popularity_txt %>%
  whisker.render(data) %>%
  cat

```

```{r}
dep_popularity_txt %>%
  whisker.render(data) %>%
  dbGetQuery(sc, .) %>%
  ggplot(
    aes(reorder(department_id, n_ord, function(x) x), 
        n_ord)) +
  geom_bar(stat="identity", fill='darkorange') +
  coord_flip() + 
  scale_y_continuous(label=scales::comma) +
  xlab("Department") +
  ylab("Number of products") +
  theme_minimal()
```

## More on orders (Window Functions)

```{r}
dbGetQuery(sc, 
"
SELECT days_since_prior_order
,   COUNT(1) AS N
FROM orders_tbl
GROUP BY days_since_prior_order
") 
```


```{r}
dbGetQuery(sc, 
"SELECT * 
FROM orders_tbl
WHERE user_id <= 10
")
```
```{r}
dbGetQuery(sc, 
"SELECT user_id
,   order_id
,   FIRST_VALUE(order_id)  OVER(partition by user_id ORDER BY order_number) AS first_order
FROM orders_tbl
WHERE user_id <= 10 
")
```

* Product that is usually ordered in numerous baskets

```{r}
"
SELECT product_id
,   AVG(products_number) AS avg_products_number
FROM (
    SELECT order_id
    ,   product_id
    ,   COUNT(product_id)  OVER(partition by order_id) AS products_number
    FROM order_products__train_tbl
) pn
GROUP BY product_id
ORDER BY avg_products_number DESC
LIMIT 100
" %>%
dbGetQuery(sc, .)
```

* Could modify the query and add product name?

```{r}
"
SELECT pn.product_id, product_name,
AVG(products_number) AS avg_products_n
FROM(
  SELECT order_id,
    product_id,
    COUNT(product_id) OVER (partition by order_id) AS products_number
    FROM order_products__train_tbl
) pn LEFT JOIN products_tbl t2 ON pn.product_id=t2.product_id
GROUP BY pn.product_id, product_name
ORDER BY avg_products_n DESC
LIMIT 20
" %>%
dbGetQuery(sc, .)
```

* Product that is usually ordered in small baskets

```{r}
"
SELECT product_id
,   AVG(products_number) AS avg_products_number
FROM (
    SELECT order_id
    ,   product_id
    ,   COUNT(product_id)  OVER(partition by order_id) AS products_number
    FROM order_products__train_tbl
) AS pn
GROUP BY product_id
ORDER BY avg_products_number
LIMIT 100
" %>%
dbGetQuery(sc, .)
```

* For 10 first users and all their orders find days since first order

```{r}
orders %>%
  filter(user_id <= 10) %>%
  select(user_id, order_number, days_since_prior_order) %>%
  arrange(user_id, order_number) %>%
  mutate(days_since_first_order = cumsum(days_since_prior_order)) %>%
  mutate(days_since_first_order=ifelse(is.null(days_since_prior_order), 0, days_since_first_order))
```

```{r}
"
SELECT user_id, order_number, COALESCE(days_since_prior_order,0) AS days_since_prior_order, COALESCE(SUM(days_since_prior_order) OVER
(PARTITION BY user_id ORDER BY order_number ROWS BETWEEN UNBOUNDED PRECEDING AND CURRENT ROW),0) AS days_since_first_order
FROM orders_tbl
WHERE user_id <= 10
"%>%
dbGetQuery(sc, .)
```











